Examine the efficacy of ascaroside #18 in controling Escherichia Coli on alfalfa and fenugreek seeds and sprouts

### 2024-02-24 R/analysis-code/exploratory-analysis-v2.qmd
mistakes fixed

---
title: "Project data cleaning"
author: "Xueyan Hu"
date: "2023-02-22"
output: html_document
---

@@

Also note that while here I split cleaning and exploring, this is iterative. You saw that as part of the processing, we already had to explore the data somewhat to understand how to clean it. In general, as you explore, you'll find things that need cleaning. As you clean, you can explore more. Therefore, at times it might make more sense to combine the cleaning and exploring code parts into a single R or Quarto file. Or split things in any other logical way.

As part of the exploratory analysis, you should produce plots or tables or other summary quantities for the most interesting/important quantities in your data. Depending on the total number of variables in your dataset, explore all or some of the others. Figures produced here might be histograms or density plots, correlation plots, etc. Tables might summarize your data.

Start by exploring one variable at a time. Then continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in. If your dataset is small, you can do that for all variables.

Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones.

@@

```{r, include=FALSE, cache=FALSE}
knitr::read_chunk('edacode.R')
```

@@

```{r,table1}
```

@@

# Data exploration through figures

Seed type first: bar chart

```{r,seed}
```

Then goes strain type: box plot

```{r,strain}
```

Now the effect of the treatment: bar chart with color gradient as an indicator

```{r,treatment}
```

The growth trend of the bacterial population over time: scatterplot with line

```{r,day}
```

linear model applied

```{r,modelfit}
```

# Notes

>>>>

---
title: "An example exploratory analysis script with code pulled in"
author: "Yao Lu, Andreas Handel"
date: "2023-03-13"
output: html_document
---

@@

@@

```{r, include=FALSE, cache=FALSE}
knitr::read_chunk('exploratorycode.R')
```

@@

```{r,table 1}
```

@@

# Data exploration through through fit model

```{r,ANOVA table}
```

# LSD test

```{r,LSD test}
```
# Notes

### 2024-02-24 R/analysis-code/exploratory-analysis.qmd
mistakes fixed

anovatable_file = here("results", "tables", "anovatable.rds")
saveRDS(linear_model, file = anovatable_file)
```

>>>>

anovatable_file = here("results", "tables", "anovatable.rds")
saveRDS(linear_model, file = anovatable_file)

# Fit an new ANOVA model
model1 <- aov(Selective ~ Treatment, data = mydata)

# Perform LSD test
lsd_result <- LSD.test(model1, "Treatment", alpha = 0.05)

# Display the results
print(lsd_result)

# save LSD result

lsdtable_file = here("results", "tables", "lsdtable.rds")
saveRDS(lsd_result, file = lsdtable_file)
```

### 2024-02-24 products/manuscript/Manuscript copy.qmd
mistakes fixed

The structure below is one possible setup for a data analysis project (including the course project).
For a manuscript, adjust as needed.
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format.
[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
You can switch to other formats, like html or pdf.
See [the Quarto documentation](https://quarto.org/) for other formats.

@@

## Data aquisition

@@

The detailed raw data and processed data are stored in 'data' folder.

*Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along.*

@@

*Explain anything related to your statistical analyses.*

@@

*Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project.*

>>>>

@@

## Data acquisition

@@

The detailed raw data and processed data are stored in 'data' folder.
And the code for data cleaning is in the 'processing-code' subfolder under 'R' folder.
Since the name of the categories in variable "Day" is a bit confusing, I changed it to the form of day+time point0,1,3,5,7.

And the reason why I only choose the colony count from selective media is that supportive media may have more colonies that are not specific target strain due to its ingredient feature.
So the count from selective media could be more reliable.

@@

By fitting the data with ANOVA model, the result will show whether each independent variable has main effect to the dependent variable through the P value.
And the summary table is stored in 'table' subfolder under 'result' folder.

Then for the efficacy of chemical, a LSD test is applied and the significant level will be presented for the 4 treatments.

Bar chart, box plot will show the difference of bacterial populations between 2 types of seeds, 2 types of strains, and among 4 treatments.
As well as line graph for the bacteria growth trend during sprouting process.

@@

My main interest here is the realtionship between treatment and bacterial populations.
Based on the ANOVA table, seed type, treatment and day are the variables that have significant effect on the dependent variables; while strain type is not.
And there is no significant difference between 2 replications, which means the experiment mayvbe replicable.
Furthermore, I ran a LSD test to check the significant level for each 4 treatment and got the result as follows.

Table 1.
The result of ANOVA model.

| Variable  | Df  | Sum Sq | Mean Sq | F value | Pr(F)    |
|-----------|-----|--------|---------|---------|----------|
| Seed      | 1   | 95.8   | 95.81   | 43.287  | 7.53e-10 |
| Strain    | 1   | 0.1    | 0.10    | 0.047   | 0.829    |
| Treatment | 3   | 322.3  | 107.43  | 48.538  | \< 2e-16 |
| Day       | 4   | 113.8  | 28.46   | 12.859  | 5.12e-09 |
| Rep       | 1   | 0.4    | 0.42    | 0.190   | 0.664    |
| Residuals | 149 | 329.8  | 2.21    |         |          |

Table 2.
Treatment summary table.

| Treatment type     | Mean population | Significant level |
|--------------------|:---------------:|-------------------|
| Control 2          |      4.79       | a                 |
| Control 1          |      4.49       | a                 |
| Chemical treated 2 |      3.23       | b                 |
| Chemical treated 1 |      1.18       | c                 |

The figures created are for some basic concepts about the population difference among the categories.

![](images/seed_type.png){fig-align="center" width="500"}

Figure 1.
The population difference between alfalfa and fenugreek seeds.

![](images/strain_type.png){fig-align="center" width="500"}

Figure 2.
The population difference between F strain and G strain (no significant difference).

![](images/treatment.png){fig-align="center" width="500"}

Figure 3.
The effect of the chemical treatment application representing by color.

![](images/timepoint.png){fig-align="center" width="500" height="450"}

Figure 4.
The growth trend of bacteria population during sprouting process.
